---
title: "Reading MACC and Traffic Count Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load and inspect a NetCDF file

To work with netCDF files, the package 'ncdf4' can be used (alternatively, you might want to check 'RNetCDF').
`nc_open` opens the file and shows its properties/content. 
The 'raster' package provides an easy way to work with the data. The function `stack` can be used to directly load a netcdf file into a raster stack.

```{r}
library(ncdf4)
library(raster)
nc <- nc_open("MACC/SURF_ENS_2012022400.nc")
No2 <- stack("MACC/SURF_ENS_2012022400.nc", varname="no2_conc")
plot(No2)
```

However, that might not work in all cases. Trying the same with a netcdf file I downloaded from the Macc webpage gives an error.

```{r, eval=F}
library(ncdf4)
getwd()
nc <- nc_open("MACC/SURF_ENS_2012022400.nc")
No2 <- stack("MACC/SURF_ENS_2012022400.nc", varname="no2_conc")
```


Alternative:`ncvar_get` gives an array of values of a certain variable, here 'no2_conc'. It has 4 dimensions, the third is the level (Surface, 50m 250m, 500m, 1000m, 2000m, 3000m and 5000m), the 4th is time (hours).
However, rows and columns need to be swapped. The following is probably not necessary for the MACC data provided in the sciebo folder, because `stack()` works for those!

```{r}
nc <- nc_open("MACC/SURF_ENS_2012022400.nc")
No2 <- ncvar_get(nc,"no2_conc")
No2 <- aperm(No2,c(2,1,3)) #swap x and y axis
```

To convert the array into a raster brick (with one layer for each time step, i.e. 24 layers for one day of hourly measurements), the function `brick` can be used.
When loading a netcdf file into a raster brick this way, the extent and crs need to be defined manually. Here, information from the `ncdf4` object can be used.

```{r}
library(raster)
#get only surface level as 3dimensional array and convert to RasterBrick
No2.r <- brick
plot(No2.r)
#set extent using information from netcdf file
xmin(No2.r)<-nc$dim$longitude$vals[1]-360
xmax(No2.r)<-nc$dim$longitude$vals[length(nc$dim$longitude$vals)]
ymin(No2.r)<-nc$dim$latitude$vals[length(nc$dim$latitude$vals)]
ymax(No2.r)<-nc$dim$latitude$vals[1]

#set CRS
proj4string(No2.r) <- "+proj=longlat +datum=WGS84 "


library(mapview)
plot(No2[[12]])
test <- No2[[12]]
coordinates(test)
mapview(rasterToPolygons(test))
```


## Load and inspect data from traffic intensity stations

The traffic intensity data is available as csv file containing (among others) the id of stations and the traffic counts. The location of the stations is available in a separate file (csv or shapefile) and needs to be joined with the count data.

```{r}
#read data of traffic count stations, here August 2017
counts <- read.csv("08-2017.csv", sep = ";", skipNul = TRUE)
#read locations of stations
locations <- read.csv("traffic_locations/pmed_trafico.csv", sep = ";", skipNul = TRUE)
traffic <- merge.data.frame(counts,locations,by.x="identif",by.y = "cod_cent")
class(traffic)
coordinates(traffic)  <- c("st_x","st_y")
class(traffic)
```

By specifying the coordinates, the data frame is transformed into a `SpapialPointsDataFrame`. However, we have multiple measurements at each location (every 15 minutes), so we want a format that facilitates handling the time dimension. We can convert the `SpatialPointsDataFrame` into a `SpaceTimeFullDataFrame`.

```{r}
library(spacetime)
#convert time value (fecha) from factor to POSIXct
traffic$time <- as.POSIXct(traffic$fecha) 
#create spacetime full data frame (via spacetime irregular data frame, to cope with NAs)
trafficSTI <- STIDF(SpatialPoints(coords = coordinates(traffic)), traffic$time, data = traffic@data)
trafficSTF <- as(trafficSTI, "STFDF")
```

This facilitates e.g. selection in space and time
```{r}
trafficSTF[1,1:12,"intensidad"] #select first spatial feature (station), time instances 1 to 12, and variable "intensidad"
trafficSTF[1,"2017-08-01 01:15:00::2017-08-01 02:00:00", "intensidad"] #select first spatial feature (station), time instances between 01:15:00 and 02:00:00 on 1 August 2017, and variable "intensidad"

```

You can also assign names to the rows of the `STFDF` to enable selection of stations by names or ids.

```{r}
row.names(trafficSTF@sp)<-trafficSTF@data$identif[1:length(trafficSTF@sp)] #assign rownames, here taken from coloumn "identif"
trafficSTF["05FT10PM01","2017-08-01 01:15:00::2017-08-01 02:00:00", "intensidad"] #select station 28045, time instances between 01:15:00 and 02:00:00 on 1 August 2017, and variable "intensidad"
```

We can also set the CRS of the data.
```{r}
proj4string(trafficSTF) <- "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
coordinates(trafficSTF)

mapview(rasterToPolygons(trafficSTF@sp))
  ```